Modal type theory based on the intuitionistic epistemic logic

### 2018-01-29 article.tex
added proofs

@@

i) Strong normalization for ${\bf IK}$ was proved by Kakutani for call-by-value and for call-by name \cite{ModalK} \cite{ModalK1}.

@@

  Unfold $M \twoheadrightarrow_{\beta} N$ as the sequence of one-step reduction and apply strip lemma on the every step.
\end{proof}

@@

\section{Monadic extention}

In modern Haskell \verb"Monad" class is an extention of the \verb"Applicative" class,
where monadic binding (\verb"(>>=)") is the single basic class method:

\begin{lstlisting}[language=Haskell]
  class Applicative m => Monad (m :: * -> *) where
    (>>=) :: m a -> (a -> m b) -> m b
    (>>) :: m a -> m b -> m b
    return :: a -> m a
    fail :: String -> m a
\end{lstlisting}
From a logical point of view we may consider this extention as the extention of IEL$^{-}$ with the imdepotency axiom.

\begin{defin} ``Monadic'' intuitionistic epistemic logic IEL$^{-}$

  4) ${\bf KK} A \to {\bf K}A$

  Rule: MP.
\end{defin}

Categorically \verb"Monad" class is a Kleisli triple on monoidal category, or, which is the same, strong monoidal monad:

\begin{defin} Strong monoidal monad

\begin{defin} Monadic $\lambda_{{\bf K}}$

  Monadic $\lambda_{{\bf K}}$ is a $\lambda_{{\bf K}}$ with the next additional typing rule:

  \begin{prooftree}
    \AxiomC{$\Gamma \vdash M : {\bf KK}A$}
    \RightLabel{${\bf K}^2$--e}
    \UnaryInfC{$\Gamma \vdash {\bf join \:} M : {\bf K}A$}
  \end{prooftree}
\end{defin}

\begin{defin}
\end{defin}

>>>>

3) If the derivation ends in

\begin{prooftree}
\AxiomC{$\Gamma \vdash {\bf pure \:} \vec{M} : {\bf K}\vec{A}$}
\AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
\BinaryInfC{$\Gamma \vdash {\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N : {\bf K}B$}
\end{prooftree}

So $\Gamma \vdash \vec{M} : \vec{A}$ by generation and $\Gamma \vdash N [\vec{x} := \vec{M}] : B$ by weakening and substitution.

Then we can transform this into the next derivation:

\begin{prooftree}
\AxiomC{$\Gamma \vdash N [\vec{x} := \vec{M}] : B$}
\RightLabel{${\bf K}_I$}
\UnaryInfC{$\Gamma \vdash {\bf pure \:} N [\vec{x} := \vec{M}] : {\bf K}B$}
\end{prooftree}

@@

We modify and apply Tait's technique of logical relation for modalities.
For strong normalization proof with Tait's method for simply typed lambda calculus see \cite{Pierce}.

\begin{defin} The set of strongly computable terms:
  \begin{itemize}
    \item $SC_A = \{ M : A \: | \: M \text{ is strongly normalizing} \}$ for $A \in \mathbb{T}$;
    \item $SC_{A \to N} = \{ M : A \to B \: | \: \forall N \in SC_A, M N \in SC_B \}$, for $A,B \in \mathbb{T}_{\bf K}$
    \item $SC_{{\bf K}A} = \{ M : {\bf K}A \: | \: M \text{ is strongly normalizing} \}$ for $A \in \mathbb{T}$;
    \item $SC_{{\bf K}(A \to B)} = \{ M : {\bf K}(A \to B) \: | \: \forall f \in SC_{A \to B}, \forall x \in SC_A, \forall N \in SC_{{\bf K}A}, {\bf let \: pure \:} f, x = M, N {\: \bf in \:} f x \}$ for $A,B \in \mathbb{T}_{\bf K}$.
  \end{itemize}
\end{defin}

\begin{itemize}
\item If $M \in SC_A$, then $M$ is strongly normalizing;
\item Let $M \in SC_A$ and $M \rightarrow_{\beta} N$, then $N \in SC_A$;
\item Let $N$ is non-introduced, $N \in SC_A$. Then, if $M \rightarrow_{\beta} N$, then $M \in SC_A$;
\end{itemize}
\end{lemma}

\begin{proof}
  By induction on the structure of $A$.

  1) $A \equiv {\bf K}A$, where $A \in \mathbb{T}$.

  i) Follows from the definition;

  ii) Immediately;

  iii) Let $N$ is non-introduced and $N \in SC_A$, such that $M \rightarrow_{\beta} N$.
  Any reduction path $M \rightarrow_{\beta} \dots$ passes through $M \rightarrow_{\beta} N$.

  $N$ is strongly normalizing, so $M$ too.

  2) $A \equiv {\bf K}(B \to C)$

  i) Suppose $M \in SC_{{\bf K}(B \to C)}$. Let $N \in SC_{{\bf K}B}$. So ${\bf let \: pure \:} f, x = M, N {\: \bf in \: } f x \in SC_{{\bf K}C}$.

  So $M$ is strongly normalizing, since ${\bf let \: pure \:} f, x = M, N {\: \bf in \: } f x$ is strongly normalizing.

  ii) Let $M_1 \in SC_{{\bf K}(B \to C)}$ and $M_1 \rightarrow_{\beta} M_2$. Fix $N \in SC_{{\bf K}B}$.

  Then ${\bf let \: pure \:} f, x = M_1, N {\: \bf in \: } f x \in SC_{{\bf K}C}$.

  Hence, ${\bf let \: pure \:} f, x = M_1, N {\: \bf in \: } f x \in SC_{{\bf K}C} \rightarrow_{\beta} {\bf let \: pure \:} f, x = M_2, N {\: \bf in \: } f x$.

  So ${\bf let \: pure \:} f, x = M_2, N {\: \bf in \: } f x \in SC_{{\bf K}C}$. Then $M_2 \in SC_{{\bf K}(B \to C)}$.

  iii) Let $M_2$ be non-introduced, $M_2 \in SC_{{\bf K}(B \to C)}$ and $M_1 \rightarrow_{\beta} M_2$.

  Let $N \in SC_{{\bf K}B}$.
\end{proof}

@@

  Unfold $M \twoheadrightarrow_{\beta} N$ as the sequence of one-step reductions and apply strip lemma on the every step.
\end{proof}

@@
